Malacology: A Programmable Storage System

### 2016-10-13 paper/paper.tex
paper: add Noah's problem comments
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Building storage stacks from the ground-up for a specific purpose results in
the best performance. For example, HDFS was designed specifically to serve
Hadoop's jobs, and uses techniques like relaxing POSIX constraints to achieve
the best performance~\cite{CITEME}. Alternatively, general-purpose storage
stacks are built with the flexiblity to serve many applications by exposing
tunable parameters to control low level implementation details like block size.
Unfortunately, users want more control from the general-purpose storage stacks
without going as far as building their storage system from the ground up.

@@

To demonstrate this trend in storage systems, we examine Ceph's
programmability. Something of a storage swiss army knife, Ceph supports file,
block, and object interfaces simultaneously in a single
cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
Distributed Object Storage (RADOS) system is a cluster of storage devices
(OSDs) that provide Ceph with data durability and integrity using replication,
erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already has some
programmability; the OSDs support the implementation of domain-specific logical
object interfaces by composing existing interfaces. These interfaces are
written in C++ and are statically loaded into the system.

@@

The Ceph community provides empirical evidence that developers embrace
programmable storage. Figure~\ref{fig:obj-int-dev-growth} shows a dramatic
growth in the production use of domain-specific interfaces in the Ceph
community since 2010.  What is most remarkable is that this trend contradicts
the notion that API changes are a burden for users.  Rather it appears that a
gap in existing interfaces are being addressed through programmability. In
fact, Table~\ref{table:objclasses} categorizes existing interfaces and we
clearly see a trend towards reusable services.

@@

The takeway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
already trying to achieve programmability because their needs, whether they be
related to performance, availability, consistency, correctness etc., are not
satisfied by the underlying storage system. The popularity of the custom object
interface facility of Ceph could be due to a number of reasons, such as the
default algorithms/tunables of the storage system are insufficient for the
application's performance goals, programmers want to exploit
application-specific semantics, and/or programmers know how to manage resources
to improve performance. Their solution is a way to work around the
traditionally rigid storage APIs because custome object interfaces give
programmers the ability to tell the storage system about their application: if
it is CPU or IO bound, if it has locality, if its size has the potential to
overload a single proxy node, etc.  The programmers know what the problem is
and how to solve it, but until the ability to modify object interfaces, they
had no way to express to the storage system how to handle their data.

Our approach is to expose more of the commonly used, code-hardened subsystems
of the underlying storage system as interfaces. The intent is that these
interfaces, which can be as simple as a redirection to the persistent data
store or as complicated as a strongly consistent directly service, should be
used and re-used in many contexts to implement a wide range of services. By
making programmability a `feature', rather than a `hack' or `workaround', we
help standardize a development process that is largely ad-hoc.

>>>>

Building storage stacks from the ground-up for a specific purpose results in
the best performance. For example, HDFS was designed specifically to serve
Hadoop's jobs, and uses techniques like exposing data locality and relaxing POSIX constraints to achieve
the best performance~\cite{CITEME}. Alternatively, general-purpose storage
stacks are built with the flexiblity to serve many applications by providing a
variety of interfaces and tunable parameters. Unfortunately, managing competing
forces in these systems is difficult and users
want more control from the general-purpose storage stacks
without going as far as building their storage system from the ground up.

@@

To demonstrate this trend in storage systems we examine the state of
programmability in Ceph. Something of a storage swiss army knife, Ceph simultaneously supports file,
block, and object interfaces on a single
cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
Distributed Object Storage (RADOS) system is a cluster of storage devices
(OSDs) that provide Ceph with data durability and integrity using replication,
erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already has some
programmability; the OSDs support domain-specific object interfaces that are
implemented by composing existing low-level storage interfaces that execute
atomically. These interfaces are
written in C++ and are statically loaded into the system.

@@

The Ceph community provides empirical evidence that developers embrace
programmable storage. Figure~\ref{fig:obj-int-dev-growth} shows a dramatic
growth in the production use of domain-specific interfaces in the Ceph
community since 2010. In that figure, classes are functional groupings of methods on storage
objects.  What is most remarkable is that this trend contradicts the notion
that API changes are a burden for users.  Rather it appears that a gap in
existing interfaces are being addressed through programmability. In fact,
Table~\ref{table:objclasses} categorizes existing interfaces and we clearly see
a trend towards reusable services.

@@

The takeway from Figure~\ref{fig:obj-int-dev-growth} is that programmers are
already trying to use programmability because their needs, whether they be
related to performance, availability, consistency, convenience, etc., are not
satisfied by the existing default set of interfaces. The popularity of the custom object
interface facility of Ceph could be due to a number of reasons, such as the
default algorithms/tunables of the storage system are insufficient for the
application's performance goals, programmers want to exploit
application-specific semantics, and/or programmers know how to manage resources
to improve performance. Their solution is a way to work around the
traditionally rigid storage APIs because custom object interfaces give
programmers the ability to tell the storage system about their application: if
it is CPU or IO bound, if it has locality, if its size has the potential to
overload a single proxy node, etc.  The programmers know what the problem is
and how to solve it, but until the ability to modify object interfaces, they
had no way to express to the storage system how to handle their data.

Our approach is to expose more of the commonly used, code-hardened subsystems
of the underlying storage system as interfaces. The intent is that these
interfaces, which can be as simple as a redirection to the persistent data
store or as complicated as a strongly consistent directory service, should be
used and re-used in many contexts to implement a wide range of services. By
making programmability a `feature', rather than a `hack' or `workaround', we
help standardize a development process that is largely ad-hoc.
